An empirical study of fixation for strategies in the Iterated Prisoner's Dilemma

### 2017-04-28 tex/main.tex
Add some discussion and references.

The Moran process was initially introduced in \cite{Moran1957} in a genetic
setting. It has since been used in a variety of settings including the
understanding of the spread of cooperative behaviour. However, as stated before,
these mainly consider non sophisticated strategies. Some work has looked at
evolutionary stability of strategies within the Prisoner's Dilemma \cite{Li2014}
but this is not done in the more widely used setting of the Moran process but in
terms of infinite population stability. In \cite{Baek2016} Moran processes are
looked at in a theoretic framework for a small subset of strategies. In
\cite{Lee2015} machine learning techniques are used to train a strategy capable
of resisting invasion and also invade any memory one strategy. Recent work
\cite{Hilbe2017} has investigated the effect of memory length on strategy
performance and the emergence of cooperation but this is not done in Moran
process context and only considers specific cases of memory 2 strategies.
% TODO Is this OK? We should add more.

@@

The main fixation probabilities of interest are \(x_1\) and \(x_{N-1}\), these
reflect a strategy's ability to invade or resist invasion.
% TODO Perhaps mention that a lot of the research in to training fixation of
% strategies assumes N=2? Is this true?
For \(N=2\) these two cases coincide. Figure~\ref{fig:fixation_heatmap_2} shows
all pairwise fixation probabilities for strategies on the vertical column when
being matched against probabilities on the horizontal column. This is summarised
in Figure~\ref{fig:boxplot_2} and Table~\ref{tbl:summary_top_2} It can be seen
that, as described in in~\cite{Press2012}, zero determinant strategies perform
well. The strategy with infinite memory \texttt{Collective} is a strategy with a
handshake which implies that it will perform well against itself.. There is also
one of the trained finite state machines in the top five.  The remaining
strategy ``Feld" is the corresponding strategy submitted to the first tournament
by Axelrod~\cite{Axelrod1980a}. All these strategies are stochastic.

\begin{figure}[!hbtp]
    \centering
    \includegraphics[width=.5\textwidth]{../img/fixation_heatmap_2_invade.pdf}
    \caption{The pairwise fixation probabilities for \(N=2\)}
    \label{fig:fixation_heatmap_2}
\end{figure}

\begin{figure}[!hbtp]
    \centering
    \includegraphics[height=.5\textheight]{../img/boxplot_2_invade.pdf}
    \caption{The median fixation probabilities for \(N=2\)}
    \label{fig:boxplot_2}
\end{figure}

@@

    %TODO Replace N=12 with N=14
    \end{subfigure}%

@@

% TODO Write a little conclusion about invaders and introduce next section.

>>>>

The Moran process was initially introduced in \cite{Moran1957} in a genetic
setting. It has since been used in a variety of settings including the
understanding of the spread of cooperative behaviour. However, as stated before,
these mainly consider non sophisticated strategies. Some work has looked at
evolutionary stability of strategies within the Prisoner's Dilemma \cite{Li2014}
but this is not done in the more widely used setting of the Moran process but in
terms of infinite population stability. In \cite{Baek2016} Moran processes are
looked at in a theoretic framework for a small subset of strategies. In
\cite{Lee2015} machine learning techniques are used to train a strategy capable
of resisting invasion and also invade any memory one strategy. Recent work
\cite{Hilbe2017} has investigated the effect of memory length on strategy
performance and the emergence of cooperation but this is not done in Moran
process context and only considers specific cases of memory 2 strategies.
% TODO Add discussion about ZD strategies and discuss how this corresponds to
% case of N = 2. Include pointing at \cite{Adami2013} which discusses how ZD
% strategies are not fantastic in population dynamics.
% Add \cite{Li2014}

@@

The main fixation probabilities of interest are \(x_1\) and \(x_{N-1}\), these
reflect a strategy's ability to invade or resist invasion.
% TODO Perhaps mention that a lot of the research in to training fixation of
% strategies assumes N=2? Is this true?
For \(N=2\) these two cases coincide. Figure~\ref{fig:fixation_heatmap_2} shows
all pairwise fixation probabilities for strategies on the vertical column when
being matched against probabilities on the horizontal column. This is summarised
in Figure~\ref{fig:boxplot_2} and Table~\ref{tbl:summary_top_2}.

\begin{enumerate}
    \item The top strategy is an extortionate Zero determinant
        strategy~\cite{Press2012} with parameters \(l=1\) and \(s=1/4\).
    \item The Collective strategy has a simple handshake mechanism (a
        cooperation followed by a defection on the first move). It then either
        cooperates or defects for all rounds~\cite{Li2009}.  This strategy was
        specifically designed for Evolutionary processes so it is perhaps also
        not surprising that it does well here.
    \item The finite state machine stragegy % TODO Perhaps refer to earlier
        % discussion in the methodology section about these FSM strategies
    \item The Feld strategy is the corresponding strategy submitted to Axelrod's
        first tournament~\cite{Axelrod1980a}: it punishes defections but
        otherwise defects with a random probability that decays over time.
    \item The final strategy in the top five is another extortionate Zero
        determinant strategy~\cite{Press2012} with parameters \(l=p\).
\end{enumerate}

\begin{figure}[!hbtp]
    \begin{subfigure}[t]{.4\textwidth}
        \centering
        \includegraphics[width=\textwidth]{../img/fixation_heatmap_2_invade.pdf}
        % TODO Remove this: I don't feel it is effective.
        \caption{The pairwise fixation probabilities for \(N=2\)}
        \label{fig:fixation_heatmap_2}
    \end{subfigure}%
    ~
    \begin{subfigure}[t]{.6\textwidth}
        \centering
        \includegraphics[height=.8\textheight]{../img/boxplot_2_invade.pdf}
        \caption{The median fixation probabilities for \(N=2\)}
        \label{fig:boxplot_2}
    \end{subfigure}
\end{figure}

@@

    %TODO Replace N=12 with N=14
    \end{subfigure}%
    %TODO Remove these? I don't feel they are effective

@@

It can be seen that apart from the Collective strategy, none of the strategies
of Table~\ref{tbl:summary_top_2} perform well for \(N\in\{3, 7, 12\}\). The new
% TODO Change N=12 to N=14
high performing strategies are:

\begin{itemize}
    \item Predator, a finite state machine described in~\cite{Ashlock2006}.
    \item Prober 4, complex strategy with an initial 20 move sequence of
        cooperations and defections~\cite{prison}. This initial sequence serves
        as some kind of handshake.
    \item Remorseful Prober, a strategy that will not immediately retaliate when
        it recognises that the opponent is itself retaliating to a random
        defection~\cite{Li2011}.
    \item  Worse and worse 2: plays tit for tat for 20 moves and then defects
        with with growing probability~\cite{prison}.
    \item Tester: a strategy submitted to the second of Axelrod's tournaments
        \cite{Axelrod1980b}.
        % TODO Check if this list is still correct for N=14
\end{itemize}

As well as noting that the memory length and complexity of these strategies are
quite complex it is interesting to note that none of them are akin to memory one
strategies. Most are not stochastic.

In the next section the performance in terms of \(x_{N-1}\) will be described:
what strategies are particularly good at resisting an invasion.
